---
title: "Causal Forest Analysis Run on Rainfed Long-Term Tillage Maize Fields"
author: "Trevor Cambron, adapted from Deines et al. 2019"
date: "8/24/2021"
output: html_document
---

# Causal Forests: Rainfed Long Term Tillage, Western US Cornbelt
## Goal: Estimate the Average Treatment Effect, Average Treatmet Effect on the Treated, and Average Treatmet Effect on the Control
In this analysis, the treatment is low-intensity tillage. 
Load desired packages 
```{r}
library(tidyverse)
library(grf)
library(naniar)
library(grf)
library(RColorBrewer)
library(earth)
library(corrplot)
library(car)
library(aod)
library(sf)
library(magrittr)
library(Rcpp)
require(Rcpp)
library(here)
library(dplyr)
library(randomForest)
require(caTools)
```

Set the data folder and filename
```{r}
dataFolder <- paste0('# Insert data folder')
dataFileNameAll <- 'switchConstantMerge_switch20082020_connPxClimate4_2022-04-11.rds'
# output scratch folder for model output rdata objects
scratchFolder <- paste0('# Insert data folder')
dir.create(file.path(scratchFolder), showWarnings = FALSE)
```

Load data and clean. This involves removing removing legacy variables that are irrelevant/ incomplete and combining others
```{r}
# all cleaned data
tillYears <- readRDS(paste0(dataFolder,dataFileNameAll)) 
drop <- c('drcl','till', 'till_2', 'V1', 'pr_grow_lag', 'tc_def_jul_lag')
tillYears <- tillYears[,!(names(tillYears) %in% drop)]
tillYears <- subset(tillYears, irrigation == 0)
drop <- c('irrigation')
tillYears <- tillYears[,!(names(tillYears) %in% drop)]
# combine some variables, remove redundancies
tillYears2 <- tillYears 
tillYears2 <- subset(tillYears2, yield > -0.01)
```

Remove SCYM (yield) outliers. Rename some variables, filter some subgroups
```{r}
# clean outliers from SCYM
summary(tillYears2$yield)

cleanQ <- quantile(tillYears2$yield, probs = c(0.0001, 0.9999), na.rm=TRUE)
cleanQ

# remove outliers:
tillYearsAll <- tillYears2 %>%
  ungroup() %>%
  mutate(uniqueID = row_number())
tillYearsAll <-   subset(tillYearsAll, yield > cleanQ[1])
tillYearsAll<- subset(tillYearsAll, yield < cleanQ[2])

#uncleaned 
sc <- tillYears$fips
sc <- substr(sc, start = 1, stop = 2)
table(sc)

length(unique(tillYearsAll$pointID)) 
tillYearsAllConstant <- tillYearsAll %>%  filter(tillStatus == 'constant_high') 
tillYearsAllConstant <- tillYearsAllConstant[,!(names(tillYearsAllConstant) %in% c("year_switched"))]
tillYearsAllConstant <- na.omit(tillYearsAllConstant)
length(unique(tillYearsAllConstant$pointID)) 
tillYearsAllConstantLow <- tillYearsAll %>%  filter(tillStatus == 'constant_low') 
tillYearsAllConstantLow <- tillYearsAllConstantLow[,!(names(tillYearsAllConstantLow) %in% c("year_switched"))]
tillYearsAllConstantLow <- na.omit(tillYearsAllConstantLow)
length(unique(tillYearsAllConstantLow$pointID)) 

table(tillYearsAll$tillStatus)

# treatment converter key
binary <- data.frame(tillStatus = c('constant_high','constant_low'),
                     W = c(0,1))

tillYearsAll2 <- tillYearsAll %>%
  left_join(binary) %>%
  rename(Y = yield)
```

Get ready to run regression forest- omit NA's and pull some covariates to build a propensity score model
```{r}
# extract balanced for training
samples_per_group <- 500
set.seed(5)
drop2 <- c("year_switched")
tillYearsAll2 <- tillYearsAll2[,!(names(tillYearsAll2) %in% drop2)]
constantLess <- tillYearsAll2 %>%
  na.omit() %>%
  group_by(ted, tillStatus, year) %>%
  sample_n(., size = min(samples_per_group, n()), replace = FALSE) %>%
  ungroup() %>%
  mutate(year = as.integer(year))

table(constantLess$tillStatus)

table(constantLess$year)

table(constantLess[,c('tillStatus','year')])

table(tillYearsAll2$tillStatus)

# how many repeat points
nrow(constantLess)

sum(duplicated(constantLess$pointID))


length(unique(constantLess$pointID))


# covariates: climate normals and soil variables
X_norm <- constantLess %>%
  dplyr::select(c(contains('norm'), contains('lay1'), 'slope',
                  'sand', 'silt', 'soc', 'clay', 'pH', 'bulkDensityt', 'rootznaws'))

# treatment
W <- constantLess %>% dplyr::pull(W)

```

Run the regression forest and check it out!
```{r}
# run
W_forest <- grf::regression_forest(X_norm, W, num.trees = 2000)

W_hat <- predict(W_forest)$predictions

WvarImp <- W_forest %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(W_forest$X.orig)) %>% 
  arrange(desc(V1))
```

```{r}
WvarImp
```

Save some data that will be extracted later (aka right now)
```{r}
# save intermediates
saveRDS(constantLess, paste0(scratchFolder, '/constantLess.rds'))
saveRDS(W_forest, paste0(scratchFolder, '/W_forest.rds'))
saveRDS(W_hat, paste0(scratchFolder, '/W_hat.rds'))
saveRDS(WvarImp, paste0(scratchFolder, '/W_forest_varImp.rds'))

# re-load so chunk runs in markdown output
constantLess <- readRDS(paste0(scratchFolder, '/constantLess.rds'))
W_hat <- readRDS(paste0(scratchFolder, '/W_hat.rds'))
W <- constantLess %>% dplyr::pull(W)
p_rf = W_hat
```


```{r}
# filter it
overlap0 <- constantLess %>%
  mutate(w_hat = W_hat) %>%
  mutate(keep = case_when(w_hat < 0.05 | w_hat > 0.95 ~ 0,
                        w_hat >= 0.05 | w_hat <= 0.95  ~ 1))
overlap <- overlap0 %>% filter(keep == 1)
nrow(overlap0) - nrow(overlap)
 
nrow(overlap)

nrow(overlap)

table(overlap$tillStatus)

table(overlap$year)

table(overlap[,c('tillStatus','year')])

# compare remaining teds
unique(constantLess$ted)

unique(overlap$ted)

both <- data.frame(W_hat = W_hat,
                   type = paste0('"even" sampling (',nrow(constantLess), ' obs)')) %>%
  bind_rows(data.frame(W_hat = overlap$w_hat,
                       type = paste0('Propensity Filtered (', nrow(overlap), ' obs)')))

ggplot(both,
       aes(x = W_hat)) +
  geom_histogram() +
  geom_vline(xintercept = 0.05, col='red') +
  geom_vline(xintercept = 0.95, col = 'red') +
  facet_wrap(~type, nrow = 2) +
  theme_bw()

{plot(smooth.spline(p_rf, W, df = 4))
  abline(0, 1)}

{plot(smooth.spline(overlap$w_hat, overlap$W, df = 4))
  abline(0, 1)}

# save overlap data!
saveRDS(overlap, paste0(scratchFolder, '/overlap.rds'))

overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))
```

Pick variables for MARS^TM (Earth) modelling.
Note that several variables such as lat, long, tmin, etc. seem to be arbitrarily removed, but these have actually been found to highly correlated to variables that either have a greater effect on yield or are much more interesting, so its not helpful to include them.
Other variables are quirks in the data or not super helpful for the analysis (i.e. climate norms)
```{r}
# pick subset of variables by dropping stuff: candidate 1
X_c1 <- overlap %>%
  dplyr::select(-c(pointID, latitude, longitude, tmin_jul, tmin_apr, tmax_jul, ppt_apr, fips, meanTemp_may, constTillStatus, count, nObs, W,tillStatus, uniqueID,silt,clay, nccpi3corn,
            contains('norm'), contains('ted')))
X_c1 <- X_c1[,!(names(X_c1) %in% c('mukey', 'w_hat'))]

# mars for var selection
set.seed(5)
marsAll <- earth::earth(Y ~ ., data = X_c1, degree = 2)
summary(marsAll)
ev <- evimp(marsAll) # estimate variable importance
plot(ev)
ev

saveRDS(marsAll, paste0(scratchFolder, '/mars_c1.rds'))


# variables from mars
marsAll <- readRDS(paste0(scratchFolder, '/mars_c1.rds'))
summary(marsAll)
ev <- evimp(marsAll) # estimate variable importance
plot(ev)
ev

envvars <- rownames(ev)

# This approach just adds JJA solar radiation, keeping the other variables limited to those identified by MARS
varsToKeep = c(envvars, 
                    'tc_def_jun','srad_jja', 
                    'slope',  'vpd_july', 
                    'tc_soilm_apr', 'tc_def_jun','sand','soc', 'rootznaws', 'pr_early')
```

Run a regression forest 
```{r}
X <- overlap %>%
  dplyr::select(all_of(varsToKeep))

corr <- cor(X)
corrplot(corr, type = 'upper')

# outcome
Y <- overlap %>% dplyr::pull(Y)

Y_forest <- grf::regression_forest(X, Y, num.trees = 2000)
```

Check the variable importance in predicting yield
```{r}
YvarImp <- Y_forest %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(Y_forest$X.orig)) %>% 
  arrange(desc(V1))
YvarImp
```

```{r}
Y_hat <- predict(Y_forest)$predictions

Ychecker <- data.frame(Y = Y,
                       Y_hat = Y_hat,
                       year = overlap$year)
Yagreement <- lm(Y~Y_hat)
summary(Yagreement)

ggplot(Ychecker, aes(x=Y, y = Y_hat)) +
  geom_hex() +
  scale_fill_gradientn(colors = brewer.pal(9,'YlGnBu')) +
  geom_smooth(method= 'lm') +
  coord_equal(xlim=c(0,18), ylim=c(0,18)) + 
  geom_abline(slope = 1, intercept = 0, linetype='dashed') + 
  facet_wrap(~year) +
  theme_bw()

saveRDS(Y_hat, paste0(scratchFolder, '/Y_hat.rds'))
saveRDS(Y_forest, paste0(scratchFolder, '/Y_forest.rds'))
saveRDS(YvarImp, paste0(scratchFolder, '/Y_forest_varImp.rds'))
```

Get ready to run causal forest, optionally reload variables
```{r}
# Causal Forest
# reload overlap
#overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))
#Y_hat <- readRDS(paste0(scratchFolder, '/Y_hat.rds'))

# reload mars variables
#marsAll <- readRDS(paste0(scratchFolder, '/mars_c1.rds'))
ev <- evimp(marsAll) # estimate variable importance
envvars <- rownames(ev)

# treatment
W <- overlap %>% dplyr::pull(W)
w_hat <- overlap$w_hat

# outcome
Y <- overlap %>% dplyr::pull(Y)

X_norm <- overlap %>%
  dplyr::select(c(contains('norm'), contains('lay1'),'rootznaws', 'slope'))

# make covariate list; 
# add in important propensity variables, and things i like (GDD), sherries impt var (tcsoilmapr)
varsToKeep_main = c(envvars, 
                    'tc_def_jun','srad_jja', 
                    'slope',
                    'tc_soilm_apr', 'tc_def_jun','sand','soc', 'rootznaws', 'pr_early')

Xmain <- overlap %>%
  dplyr::select(all_of(varsToKeep_main))

corr <- cor(Xmain )
corrplot(corr, type = 'upper')
```

Run the causal forests with 2000 trees for robust results 
```{r}
cf <- grf::causal_forest(Xmain, Y, W, Y_hat, w_hat, num.trees = 2000)
```

Save variable importance 
```{r}
varImpMain <- cf %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(cf$X.orig)) %>% 
  arrange(desc(V1))


saveRDS(cf, paste0(scratchFolder, '/cf_xc1.rds'))
saveRDS(varImpMain, paste0(scratchFolder, '/cf_xc1_varImp.rds'))
overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))
```

Look at some Causal forest results 
```{r}
# marsAll <- readRDS(paste0(scratchFolder, '/mars_c1.rds'))


#W_forest <- readRDS(paste0(scratchFolder, '/W_forest.rds'))
#W_hat <- readRDS(paste0(scratchFolder, '/W_hat.rds'))
W_hat <- overlap %>% dplyr::pull(w_hat)
WvarImp <- readRDS(paste0(scratchFolder, '/W_forest_varImp.rds'))

#Y_forest <- readRDS(paste0(scratchFolder, '/Y_forest.rds'))
Y_hat <- readRDS(paste0(scratchFolder, '/Y_hat.rds'))
YvarImp <- readRDS(paste0(scratchFolder, '/Y_forest_varImp.rds'))

cf <- readRDS(paste0(scratchFolder, '/cf_xc1.rds'))
varImp <- readRDS(paste0(scratchFolder, '/cf_xc1_varImp.rds'))

# covariates: climate normals and soil variables
X_norm <- overlap %>%
  dplyr::select(-c(contains('norm'), contains('lay1'), 'rootznaws', 'slope'))

# treatment
W <- overlap %>% dplyr::pull(W)
Y <- overlap %>% dplyr::pull(Y)

constantLess <- readRDS(paste0(scratchFolder, '/constantLess.rds'))
allWhat <- readRDS(paste0(scratchFolder, '/W_hat.rds'))

# process results ------------------------------
# add predicted treatment effects to data 
constOob_predict <- predict(cf)

# add hats to original data frame
overlapW0 <- overlap %>%
  bind_cols(data.frame(Y_hat = Y_hat))

colNormals <- names(X_norm)
colVars <- varImp %>% dplyr::pull(variable)

overlapW <- overlapW0 %>%
  mutate_at(colNormals, list(Q4n = ~ntile(., 4),
                             Q5n = ~ntile(., 5))) %>%
  mutate_at(colVars, list(Q4 = ~ntile(., 4),
                          Q5 = ~ntile(., 5)))

overlapTau <- overlapW %>%
  bind_cols(constOob_predict) 
```

Find the average treatment effect of the control, representing the effect of giving treatment (low till) to subjects not likely to recieve it
```{r}
atec_cf_aipw = average_treatment_effect(cf, target.sample = "control")
tauhat_rf_aipw = c(ATE=atec_cf_aipw["estimate"],
                   lower_ci=atec_cf_aipw["estimate"] - 1.96 * atec_cf_aipw["std.err"],
                   upper_ci=atec_cf_aipw["estimate"] + 1.96 * atec_cf_aipw["std.err"])
print("Tons/ hectare")
tauhat_rf_aipw
meanYield <- mean(overlapTau$Y)
meanYield
tauhat_rf_aipw_percent <- tauhat_rf_aipw/meanYield * 100
print("TPercent of Yield")
tauhat_rf_aipw_percent
```

Find average treatment effect of the treated, representing the effect of witholding treatment from sujects likely to recieve treatment
Perhaps a less interesting question than the ATEC, but you can be the judge of that
```{r}
print("AVERAGE TREATMENT EFFECT OF THE TREATED")
atet_cf_aipw = average_treatment_effect(cf, target.sample = "treated")
tauhat_rf_aipw_t = c(ATE=atet_cf_aipw["estimate"],
                   lower_ci=atet_cf_aipw["estimate"] - 1.96 * atet_cf_aipw["std.err"],
                   upper_ci=atet_cf_aipw["estimate"] + 1.96 * atet_cf_aipw["std.err"])
print("Tons/ hectatre")
tauhat_rf_aipw_t
meanYield <- mean(overlapTau$Y)
print("Percent of Yield")
meanYield
tauhat_rf_aipw_percent_t <- tauhat_rf_aipw_t/meanYield * 100
tauhat_rf_aipw_percent_t
```

Now find average treatment effect (ATE). This is run on all of the propensity score filtered data 
```{r}
ate_cf_aipw = average_treatment_effect(cf)
ate_cf_aipw = average_treatment_effect(cf, target.sample = 'all')
tauhat_rf_aipw = c(ATE=ate_cf_aipw["estimate"],
                   lower_ci=ate_cf_aipw["estimate"] - 1.96 * ate_cf_aipw["std.err"],
                   upper_ci=ate_cf_aipw["estimate"] + 1.96 * ate_cf_aipw["std.err"])
tauhat_rf_aipw
```
Now find average treatment effect on the overlap (ATO). This is run on all of the propensity score filtered data 
```{r}
ate_cf_aipw = average_treatment_effect(cf)
ate_cf_aipw = average_treatment_effect(cf, target.sample = 'overlap')
tauhat_rf_aipw = c(ATE=ate_cf_aipw["estimate"],
                   lower_ci=ate_cf_aipw["estimate"] - 1.96 * ate_cf_aipw["std.err"],
                   upper_ci=ate_cf_aipw["estimate"] + 1.96 * ate_cf_aipw["std.err"])
tauhat_rf_aipw
```
```{r}

# check propensity scores
p_rf = W_hat
hist(p_rf)

{plot(smooth.spline(p_rf, W, df = 4))
  abline(0, 1)}

both <- data.frame(slope = constantLess$slope,
                   type = '"Even" Sampled',
                   class = as.character(constantLess$W),
                   stringsAsFactors = FALSE) %>%
  bind_rows(data.frame(slope = overlap$slope,
                       type = 'Propensity Filtered',
                       class = as.character(overlap$W),
                       stringsAsFactors = FALSE))

ggplot(both,
       aes(x = slope, fill = class, group = class)) +
  geom_histogram() +
  facet_wrap(~type) +
  theme_bw()

nrow(constantLess)

nrow(overlap)

YvarImp 

Ychecker <- data.frame(Y = Y,
                       Y_hat = Y_hat,
                       year = overlap$year)
Yagreement <- lm(Y~Y_hat)
summary(Yagreement)

# yield year by year
YcheckerLong <- Ychecker %>%
  tidyr::gather(., key = variable, value = value, Y:Y_hat)


ggplot(YcheckerLong, aes(x=as.factor(year), y = value, group = interaction(year,variable), 
                         fill = variable)) +
  geom_boxplot() +
  theme_bw() +
  ylab('Yield (t/ha)') +
  xlab('Year')

# This is the causal forest variable importance
varImp
```

Find average treatment effect in terms of percent of yield 
```{r}
ate_cf_aipw = average_treatment_effect(cf, target.sample = "all")
tauhat_rf_aipw = c(ATE=ate_cf_aipw["estimate"],
                   lower_ci=ate_cf_aipw["estimate"] - 1.96 * ate_cf_aipw["std.err"],
                   upper_ci=ate_cf_aipw["estimate"] + 1.96 * ate_cf_aipw["std.err"])
tauhat_rf_aipw

meanYield <- mean(overlapTau$Y)
meanYield
tauhat_rf_aipw_percent <- tauhat_rf_aipw/meanYield * 100
tauhat_rf_aipw_percent
```

Plot some data 
```{r}
ggplot(overlapTau,
       aes(x = predictions)) +
  geom_histogram() +
  xlim(-1,2)+
  xlab('CATE') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = ate_cf_aipw["control"], col = 'red') +
  theme_bw()

ggplot(overlapTau,
       aes(x = year, y = predictions, group = year)) +
  geom_boxplot() +
  #facet_wrap(~W) +
  theme_bw() + ggtitle('oob predicted treatment effects')

# run best linear predictor analysis
calibration <- test_calibration(cf)
calibration

causalVars <- overlap %>% dplyr::select(varImp$variable)

corr <- cor(causalVars)
corrplot(corr, type = 'upper')
```
