---
title: "Rainfed Figures R"
author: Trevor Cambron
output: html_document
---
This script creates figures including the map of yield impacts, ATE/C/T boxplots, and the IQR boxplots

First load necessary packages
```{r}
library(ggpubr)
library(tidyverse)
library(grf)
library(sf)


library(raster)
library(latticeExtra)
library(RColorBrewer)

library(here)
#library(rgdal)
library(reshape)
sessionInfo()
```

Set folder destinations, import data 
```{r}
# scratch folder for model rdata objects
scratchFolder <- '# Insert outputs/ results folder from previous scripts'

# gis directory
gisFolder <- paste0(here::here(),'#Insert Spatial Data folder')

# dataset
overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))

Y_hat <- readRDS(paste0(scratchFolder, '/Y_hat.rds'))
cf <- readRDS(paste0(scratchFolder, '/cf_xc1.rds'))
varImp <- readRDS(paste0(scratchFolder, '/cf_xc1_varImp.rds'))

# process results ------------------------------
# add predicted treatment effects to data 
constOob_predict <- predict(cf)

# add hats to original data frame
overlapW0 <- overlap %>%
  bind_cols(data.frame(Y_hat = Y_hat))

colVars <- varImp %>% dplyr::pull(variable)

overlapW <- overlapW0 %>%
  mutate_at(colVars, list(Q4 = ~ntile(., 4),
                          Q5 = ~ntile(., 5)))

overlapTau <- overlapW %>%
  bind_cols(constOob_predict) 

ate_cf_aipw = average_treatment_effect(cf, target.sample= 'all')
tauhat_rf_aipw = c(ATE=ate_cf_aipw["estimate"],
                   lower_ci=ate_cf_aipw["estimate"] - 1.96 * ate_cf_aipw["std.err"],
                   upper_ci=ate_cf_aipw["estimate"] + 1.96 * ate_cf_aipw["std.err"])
tauhat_rf_aipw

# as a percentage of yield
meanYield <- mean(overlapTau$Y)
meanYield

tauhat_rf_aipw_percent <- tauhat_rf_aipw/meanYield * 100
tauhat_rf_aipw_percent

atec_cf_aipw = average_treatment_effect(cf, target.sample= 'control')
tauhat_rf_atec = c(ATE=atec_cf_aipw["estimate"],
                   lower_ci=atec_cf_aipw["estimate"] - 1.96 * atec_cf_aipw["std.err"],
                   upper_ci=atec_cf_aipw["estimate"] + 1.96 * atec_cf_aipw["std.err"])
tauhat_rf_atec

# as a percentage of yield
meanYield <- mean(overlapTau$Y)
meanYield

tauhat_rf_atec_percent <- tauhat_rf_atec/meanYield * 100
tauhat_rf_atec_percent

atet_cf_aipw = average_treatment_effect(cf, target.sample= 'treated')
tauhat_rf_atet = c(ATE=atet_cf_aipw["estimate"],
                   lower_ci=atet_cf_aipw["estimate"] - 1.96 * atet_cf_aipw["std.err"],
                   upper_ci=atet_cf_aipw["estimate"] + 1.96 * atet_cf_aipw["std.err"])
tauhat_rf_atet

# as a percentage of yield
meanYield <- mean(overlapTau$Y)
meanYield

tauhat_rf_atet_percent <- tauhat_rf_atet/meanYield * 100
tauhat_rf_atet_percent

# sampled in analysis
nrow(overlapTau)

###REVIEW
panelBackground <- 'gray30'
stateLines <- 'gray10'
stateFill <- 'gray50'

colorbreaks <- 25

# scale range maize
palMin <- -25
palMax <- 25

# blue and red color ramps pulled from color brewer (no white)
redRamp2 <- colorRampPalette(c('#fddbc7','#f4a582','#d6604d','#b2182b','#67001f'))
blueRamp2 <- colorRampPalette(rev(c('#d1e5f0','#92c5de','#4393c3','#2166ac','#08519c')))#'#053061')))

biPal <- c(blueRamp2(colorbreaks/5*2),redRamp2(colorbreaks/5*3))
```

Plot the results
```{r}
# state outlines
states <- read_sf('# sptial data folder /s_11au16.shp') %>%
  st_transform(5070) 
statesStudy <- states %>%  filter(STATE %in%             c('SD','ND','NE','KS','MN','IA','MO','IL','WI','IN','MI','OH', 'KY', 'CO','WY', 'MT','OK','TX','NM', 'AR'))

states_spdf <- as(statesStudy, 'Spatial')
statesPlot <- as(states, 'Spatial')

# raster map
aeaProj <- '+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=1,1,-1,0,0,0,0 +units=m +datum=NAD83 +no_defs'
states_spdf <- spTransform(states_spdf, CRS(aeaProj))
corners <- bbox(states_spdf)
grid <- raster(crs = aeaProj, res = 5000,
                     extent(corners[1,1],     #xmin
                            corners[1,2],      #xmax
                            corners[2,1],       #ymin
                            corners[2,2]))      #ymax
grid[] <- 1:ncell(grid)
gridmask <- mask(grid, states_spdf)
#plot(gridmask)
#plot(states_spdf, add = TRUE)
# polygonize
gridPoly <- rasterToPolygons(gridmask)
# convert to sf
grid_sf <- st_as_sf(gridPoly) %>% st_transform(4326) #%>% rename(cellID = layer)
cates <- overlapTau %>%
  mutate(tauPercents = predictions/meanYield*100)
tau_sf <- st_as_sf(cates, coords = c('longitude','latitude'), crs = 4326)
colorbreaks <- 25
# scale range maize
palMin <- -25
palMax <- 25
# blue and red color ramps pulled from color brewer (no white)
redRamp2 <- colorRampPalette(c('#fddbc7','#f4a582','#d6604d','#b2182b','#67001f'))
blueRamp2 <- colorRampPalette(rev(c('#d1e5f0','#92c5de','#4393c3','#2166ac','#08519c')))#'#053061')))
biPal <- c(blueRamp2(colorbreaks/2),redRamp2(colorbreaks/2))
```
# map vis parameters
for consistent mapping across figures
you will need to adjust the mapXLim and mapYLim
```{r mapVisParams}
panelBackground <- 'lightgrey'
stateLines <- 'black'
stateFill <- 'gray'
mapXLim <- c(-800000,180000)
mapYLim <- c(1470000,2930000)
```
## Corn
```{r}
# add grid id to point taus
tau_sf2 <- tau_sf %>%
  dplyr::select(c(year, tauPercents, predictions)) %>%
  st_join(grid_sf) 
gridTaus0 <- tau_sf2 %>%
  tidyr::drop_na() %>%
  group_by(layer) %>%
  summarize(meanTauP = mean(tauPercents),
            meanTau = mean(predictions))
gridTaus <- grid_sf %>%
  left_join(gridTaus0) %>%
  st_transform(5070)
taus_spdf <- as(gridTaus, 'Spatial')
tauRas <- rasterize(taus_spdf, gridmask, field = 'meanTauP', background = NA)
# writeRaster(tauRas, '/Volumes/LaCie/2021_Semi_Arid_Tillage/05.10Results/02.11APRIL_Rainfed_ATE_MAIZE_Results/tauRas' )
```
```{r taurasR, dev = c('png','pdf'), dpi = 300, bg="transparent"}
#tauRas <- raster("/Volumes/LaCie/2021_Semi_Arid_Tillage/05.10Results/02.10Rainfed_ATE_MAIZE_Results/maize_summaryEffects.tif")
spplot(tauRas, col.regions = biPal, axes = TRUE,
       at=seq(palMin, palMax, (-palMin + palMax)/colorbreaks),
       colorkey = list(space = 'bottom'),
       maxpixels = 1500000,
       xlim = mapXLim, ylim = mapYLim,
       par.settings = list(panel.background = list(col=panelBackground)), cex = 40) +
  latticeExtra::layer(sp.polygons(statesPlot, lwd=.8, col=stateLines)) +
  latticeExtra::layer(sp.polygons(statesPlot, lwd=.8, col=NA, fill = stateFill), under=T) 

# combine data
```
```{r}
cates <- overlapTau %>%
  mutate(tauPercents = predictions/meanYield*100,
         croptype = 'Maize') %>%
  dplyr::select(c(croptype, predictions, tauPercents, year)) 

# get ranges
cates %>%
  summarize(p5 = quantile(tauPercents, probs = c(0.05)),
            p95 = quantile(tauPercents, probs = c(0.95)))

ggplot(cates,
       aes(x = tauPercents, fill = ..x..)) +
  geom_histogram(breaks = seq(palMin, palMax, (-palMin + palMax)/colorbreaks),
                 color = 'gray30') +
  xlim(-25,25)+
  scale_fill_gradientn(colors = biPal) +
  scale_y_continuous(breaks = c(2000,4000,6000), labels = c(2, 4, 6)) +
  xlab('Yield Impact (%)') +
  ylab('Thousand Fields') +
  #geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent[1], col = 'red', linetype = 'longdash') +
  theme_bw() + theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.background = element_rect(fill = "black",
                                                     colour = NA),
                     plot.background = element_rect(fill = "black",
                                                    colour = NA),
                     panel.border = element_blank(),
                     axis.text = element_text(color = 'white'),
                     axis.line = element_line(color = 'white'),
                     axis.title = element_text(color = 'white'),
                     axis.ticks = element_line(color = 'white'),
                     legend.position = 'none') +
  theme(text = element_text(size = 35, face = 'bold'))

```

Now find ATEC 
```{r}
overlapTau.control <- subset(overlapTau, overlapTau$tillStatus == 'constant_high')
cates.control <- overlapTau.control %>%
  mutate(tauPercents = predictions/meanYield*100,
         croptype = 'Maize') %>%
  dplyr::select(c(croptype, predictions, tauPercents, year)) 

# get ranges
cates.control %>%
  summarize(p5 = quantile(tauPercents, probs = c(0.05)),
            p95 = quantile(tauPercents, probs = c(0.95)))
```
Now ATET 
```{r}
overlapTau.treated <- subset(overlapTau, overlapTau$tillStatus == 'constant_low')
cates.treated <- overlapTau.treated %>%
  mutate(tauPercents = predictions/meanYield*100,
         croptype = 'Maize') %>%
  dplyr::select(c(croptype, predictions, tauPercents, year)) 

# get ranges
cates.treated %>%
  summarize(p5 = quantile(tauPercents, probs = c(0.05)),
            p95 = quantile(tauPercents, probs = c(0.95)))
```

And make boxplots 
```{r}
cates$group <- 'ATE'
cates.control$group <- 'ATEC'
cates.treated$group <- 'ATET'

catesdf <- rbind(cates, cates.control, cates.treated)
ggplot(catesdf, aes(x = group, y = tauPercents))+
  geom_boxplot()

df.CI <- data.frame(Parameter = c("ATE", "ATEC", "ATET"), Estimate = c(tauhat_rf_aipw_percent[1],tauhat_rf_atec_percent[1], tauhat_rf_atet_percent[1]), Lower = c(tauhat_rf_aipw_percent[2],tauhat_rf_atec_percent[2], tauhat_rf_atet_percent[2]), Upper = c(tauhat_rf_aipw_percent[3],tauhat_rf_atec_percent[3], tauhat_rf_atet_percent[3]))

ggplot(df.CI, aes(x = Parameter, y = Estimate))+
  geom_point() +
  geom_errorbar(aes(ymin = Lower, ymax = Upper))
```

Add more vars:
```{r}
# bin cates: corn
ate <- tauhat_rf_aipw[1]
overlapTau_1 <- overlapTau %>%
  # code for positive or negative treatment effect
  mutate(cate_bin = case_when(predictions < 0 ~ 1,
                              predictions >= 0 & predictions <= ate ~ 2,
                              predictions > ate ~ 3),
         cate_ntile = ntile(predictions, 5))

```

# Then the IQR boxplots
```{r}
# variables to highlight
vars <- varImp[1:10,]$variable

hte_df <- overlapTau_1 %>%
  dplyr::select(c(pointID, year, predictions, cate_bin, vars)) %>%
  tidyr::gather(., key = variable, value = value, vars)

hte_df$variable[hte_df$variable == "tmax_apr"] <- "Apr Max T" #
hte_df$variable[hte_df$variable == "tmin_may"] <- "May Min T" #
hte_df$variable[hte_df$variable == "srad_jja"] <- "JJA SRad" #
hte_df$variable[hte_df$variable == "tmin_aug"] <- "Aug T Min" #
hte_df$variable[hte_df$variable == "soc"] <- "SOC" #
hte_df$variable[hte_df$variable == "ppt_may"] <- "May Ppt" #
hte_df$variable[hte_df$variable == "sand"] <- "Sand" #
hte_df$variable[hte_df$variable == "pr_early"] <- "Early Seas Ppt" #
hte_df$variable[hte_df$variable == "rootznaws"] <- "Rt Zn AWS" #
hte_df$variable[hte_df$variable == "tc_soilm_apr"] <- "Apr Soil M" #
hte_df$variable[hte_df$variable == "slope"] <- "Slope" #
hte_df$variable[hte_df$variable == "bulkDensityt"] <- "Bulk Density" #


# calculate IQR by bins
hte_iqr <- hte_df %>% 
  mutate(cate_bin = as.factor(cate_bin)) %>%
  dplyr::group_by(variable, cate_bin) %>%
  dplyr::summarize(mean = mean(value, na.rm=TRUE),
                   stdev = sd(value, na.rm=TRUE),
                   iqr_25 = quantile(value, probs = 0.25, na.rm=TRUE),
                   iqr_50 = quantile(value, probs = 0.5, na.rm=TRUE),
                   iqr_75 = quantile(value, probs = 0.75, na.rm=TRUE)) 

catePalette3 <- c("#86BDDA","#F6B394","#CF5246")
```

```{r hte_iqr, dev = c('png','pdf'), dpi = 300, bg="transparent"}
hte_iqr_1_6 <- ggplot(hte_iqr,
  aes(x = cate_bin, y = value, group = cate_bin, fill = cate_bin)) +
  geom_crossbar(aes(y = iqr_50, ymin = iqr_25, ymax = iqr_75)) +
  scale_fill_manual(values = catePalette3) +
  #ylab('') + xlab('') + 
  facet_wrap(~variable, scales = 'free_x',nrow = 2) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  scale_x_discrete(breaks=c("1","2","3"),
        labels=c("Negative", "Small\nPositive", "Above\nAverage")) +
  xlab("Treatment Effect Magnitude") +
  ylab("Value") +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=45, hjust=1))
  
```
Add more vars:
```{r}
# bin cates: corn
ate <- tauhat_rf_aipw[1]
overlapTau_1 <- overlapTau %>%
  # code for positive or negative treatment effect
  mutate(cate_bin = case_when(predictions < 0 ~ 1,
                              predictions >= 0 & predictions <= ate ~ 2,
                              predictions > ate ~ 3),
         cate_ntile = ntile(predictions, 5))

```

```{r}
rf_ate <- ate_cf_aipw["estimate"]
overlapTau <- overlapTau %>%
  mutate(cate_bin = case_when(predictions < 0 ~ "Negative",
                              predictions >= 0 & predictions <= rf_ate ~ "Small Positive",
                              predictions > rf_ate ~ "Above Average"))

rf.num1<- select_if(overlapTau, is.numeric)
rf.num <- rf.num1[ , which(apply(rf.num1, 2, var) != 0)]
rf.pca <- rf.num %>%
  #dplyr::select(c(vars <- varImp[1:5,]$variable))%>%
  prcomp(scale. = TRUE, center = TRUE)
```

Plot the results
```{r}
catePalette3 <- c("#CF5246","#86BDDA", "#F6B394")
```
                               

Make correlation plot
```{r}
M <- cor(overlapTau_1 %>%
          dplyr::select(c(VarImp[1:12,])))
corrplot(M, type="upper", order="hclust",
         col=brewer.pal(n=8, name="RdYlBu"))
```
