---
title: "07.10IdentifySubOptimum"
author: "Trevor Cambron"
date: "9/30/2021"
output: html_document
---

The goal of this script is to identify where conservation tillage could have the greatest effect, but is being underutilized.
```{r}
library(cluster)
library(tidyverse)
library(grf)
library(sf)
library(raster)
library(latticeExtra)
library(RColorBrewer)
library(here)
#library(rgdal)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(magrittr)
```
Start with improting a dataset that includes latitude, longitude and tillage status
```{r}
rainfedscratchFolder <- '# Insert Results Folder'
# datasets
rainfedoverlap <- readRDS(paste0(rainfedscratchFolder, '/overlap.rds'))
rainfedY_hat <- readRDS(paste0(rainfedscratchFolder, '/Y_hat.rds'))
rainfedcf <- readRDS(paste0(rainfedscratchFolder, '/cf_xc1.rds'))
rainfedvarImp <- readRDS(paste0(rainfedscratchFolder, '/cf_xc1_varImp.rds'))
fips <- readOGR('# Spatial Data folder /UScounties.shp')
states <- readOGR('# Spatial Data Folder /tl_2021_us_state.shp')
rfdf <- readRDS(' # Data Folder /switchConstantMerge_switch20082020_connPxClimate4_2022-04-11.rds')

NASS2017 <- readRDS(' # Data Folder /NASS2017County.RDS')
```
How many fields were constant high and low?
```{r}
df2020 <- subset(rfdf, year == 2020)

ir2020 <- subset(df2020, irrigation == 1)

rf2020 <- subset(df2020, irrigation == 0)

rfLT <- nrow(rf2020[grepl('low', rf2020$tillStatus),])
rfHT <- nrow(rf2020[grepl('high', rf2020$tillStatus),])



CrfLT <- nrow(rf2020[grepl('constant_low', rf2020$tillStatus),])
CrfHT <- nrow(rf2020[grepl('constant_high', rf2020$tillStatus),])

CrfHT/ (CrfHT + CrfLT)

```
Next process these raw results so we have something more usable 
```{r}
# process results ------------------------------
# add predicted treatment effects to data 
rainfedconstOob_predict <- predict(rainfedcf)

# add hats to original data frame
rainfedoverlapW0 <- rainfedoverlap %>%
  bind_cols(data.frame(Y_hat = rainfedY_hat))

rainfedcolVars <- rainfedvarImp %>% dplyr::pull(variable)

rainfedoverlapW <- rainfedoverlapW0 %>%
  mutate_at(rainfedcolVars, list(Q4 = ~ntile(., 4),
                          Q5 = ~ntile(., 5)))

rainfedoverlapTau <- rainfedoverlapW %>%
  bind_cols(rainfedconstOob_predict) 

# is this right? can remove
rainfedoverlapTau2020 <- subset(rainfedoverlapTau, rainfedoverlapTau$ year == 2020)

rf_ate_cf_aipw = average_treatment_effect(rainfedcf, target.sample= 'all')
tauhat_rf_aipw = c(ATE=rf_ate_cf_aipw["estimate"],
                   lower_ci=rf_ate_cf_aipw["estimate"] - 1.96 * rf_ate_cf_aipw["std.err"],
                   upper_ci=rf_ate_cf_aipw["estimate"] + 1.96 * rf_ate_cf_aipw["std.err"])
tauhat_rf_aipw

# as a percentage of yield
rf_meanYield <- mean(rainfedoverlapTau$Y)
rf_meanYield

tauhat_rf_aipw_percent <- tauhat_rf_aipw/rf_meanYield * 100
tauhat_rf_aipw_percent

# sampled in analysis
nrow(rainfedoverlapTau)
```


Find the proportion of fields that had a negative treatment effect
```{r}
neg_te <- nrow(subset(rainfedoverlapTau, predictions < 0))
total <- nrow(rainfedoverlapTau)
neg_te / total
```

The following are only for rainfed fields. 

# 1. Panel A, Find current % lt low till
```{r}

#---
rfpt <- NASS2017 %>%
  dplyr::select(c('fips', 'Proportion Low Till')) 

names(rfpt)[names(rfpt) == 'fips'] <- 'FIPS'
names(rfpt)[names(rfpt) == 'Proportion Low Till'] <- 'pLT'

fipstoPlota <- merge(fips, rfpt, by = 'FIPS') %>%
  subset(STATE_NAME %in% c('Kansas', 'North Dakota', 'South Dakota', 'Nebraska')) 



till.pal <-  colorRampPalette(c('#FD0101','#0800FF'))


statelineplot <- list("sp.polygons", states[which(states$STUSPS %in% c('KS','NE','SD','ND')),], lwd = 3)
spplot(fipstoPlota, "pLT",  col.regions = till.pal(16), main = "Low Till (% of fields)", sp.layout = statelineplot) 

```
# Panel B. County Yield Benefit
```{r}
rainfedoverlapTau2020$perfield_prediction <- rainfedoverlapTau2020$count*(.09)*rainfedoverlapTau2020$predictions

rainfedoverlapTau2020$perfield_yield <- rainfedoverlapTau2020$count*(.09)*rainfedoverlapTau2020$Y

rfOvCt <- subset(rainfedoverlapTau2020, tillStatus == 'constant_high')

rainfedATEC <- aggregate(perfield_prediction ~ fips, rfOvCt, FUN = sum )
rainfedY <- aggregate(perfield_yield ~ fips, rainfedoverlapTau2020, FUN = sum )

names(rainfedATEC)[names(rainfedATEC) == 'perfield_prediction'] <- 'sum_perfield_prediction'

names(rainfedY)[names(rainfedY) == 'perfield_yield'] <- 'sum_perfield_yield'

RFATEC <- merge(rainfedATEC, rainfedY, by = 'fips')

RFATEC$pct_yi_ltht_ltlt <- 100 * (RFATEC$sum_perfield_prediction / RFATEC$sum_perfield_yield)

RfYI <- RFATEC %>% dplyr::select(c('fips', 'pct_yi_ltht_ltlt'))

names(RfYI)[names(RfYI) == 'fips'] <- 'FIPS'

fipstoPlotc <- merge(fips, RfYI, by = 'FIPS')  %>%
  subset(STATE_NAME %in% c('Kansas', 'North Dakota', 'South Dakota', 'Nebraska'))


colorbreaks <- 18
# scale range maize
palMin <- -6
palMax <- 12 
# blue and red color ramps pulled from color brewer (no white)
redRamp2 <- colorRampPalette(c('#fddbc7','#f4a582','#d6604d','#b2182b','#67001f'))
blueRamp2 <- colorRampPalette(rev(c('#d1e5f0','#92c5de','#4393c3','#2166ac','#08519c')))#'#053061')))
biPal <- c(blueRamp2(colorbreaks*(1/3)),redRamp2(colorbreaks*(2.1/3)))

spplot(fipstoPlotc, 'pct_yi_ltht_ltlt', col.regions =biPal, main = "Yield Impact of Switching (%)", colorkey = list(height = 1, labels = list(seq(-6,12))), sp.layout = statelineplot) 

# How much would switching lt high till to lt low till cause?

sum(RFATEC$sum_perfield_prediction)

sum(RFATEC$sum_perfield_prediction) / sum(rainfedY$sum_perfield_yield)
```
# Part C. Identify high-priority counties
```{r}
HPc <- merge(rfpt, RfYI, by = 'FIPS')
 # Average percent low till 

propLT_WC <- (sum(NASS2017$`Conservation Tillage`) + sum(NASS2017$`No Till`))/
  (sum(NASS2017$`Conservation Tillage`) + sum(NASS2017$`No Till`) + sum(NASS2017$`Conventional Tillage`))

hhPC <- subset(HPc, `pLT` < propLT_WC)
HighPriorityCounties <- subset(hhPC, pct_yi_ltht_ltlt > 0)

HPCPlot <- HighPriorityCounties %>% dplyr::select(c('FIPS')) 
HPCPlot$TRU <- 1

fipstoPlotd <- merge(fips, HPCPlot, by = 'FIPS') %>%
  subset(STATE_NAME %in% c('Kansas', 'North Dakota', 'South Dakota', 'Nebraska'))

statelineplot <- list("sp.polygons", states[which(states$STUSPS %in% c('KS','NE','SD','ND')),], lwd = 3)
spplot(fipstoPlotd,'TRU',  sp.layout = statelineplot, col.regions = 'purple',  main = "High Priority Counties") 

```

Make 
```{r}
a <- plot(fipstoPlota['Proportion Low Till'],  pal = till.pal, main = "Low Till (% of fields)") + legend("bottomright", legend = levels(till.pal))
b <- plot(fipstoPlotc['pct_yi_ltht_ltlt'],   main = "Yield Impact of Switching (%)", pal = biPal) 
c <- plot(fipstoPlotd,'TRU',  main = "High Priority Counties", col.regions = 'purple') 

grid.arrange(a,b,c, ncol = 3)
```
